---
title: "collegiateSports"
author: "Braden Buchanan"
date: "2023-09-08"
output: 
  html_document:
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages
```{r}
library(tidyverse)
library(stringr)
library(forcats)
library(scales)
```

# Load data
```{r}
sports <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-03-29/sports.csv')
```

### Check out data
```{r}
colnames(sports)
str(sports)
summary(sports)
```

# Descriptive Statistics

### What schools generate the most revenue?
```{r}
highestRevSchools2019 <- sports %>%
  select(year, institution_name, state_cd, total_rev_menwomen, sports) %>%
  filter(year == 2019) %>%
  arrange(desc(total_rev_menwomen)) %>%
  head(n = 10)

ggplot(highestRevSchools2019, aes(x = total_rev_menwomen, y = fct_reorder(institution_name, total_rev_menwomen))) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = scales::dollar(total_rev_menwomen)), vjust = .3, hjust = 1.1, color = "white", size = 3.5) +
  scale_x_continuous(labels = scales::dollar_format(scale = 1)) +
  labs(title = "Top 10 Schools by Revenue in 2019",
       x = "Total Revenue",
       y = "Institution Name")
```

### What schools generate the least revenue?
```{r}
lowestRevSchools2019 <- sports %>%
  select(year, institution_name, state_cd, total_rev_menwomen, sports) %>%
  filter(year == 2019) %>%
  arrange(total_rev_menwomen) %>%
  distinct(institution_name, .keep_all = TRUE) %>%
  head(n=10)

ggplot(lowestRevSchools2019, aes(x = total_rev_menwomen, y = fct_reorder(institution_name, total_rev_menwomen))) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = scales::dollar(total_rev_menwomen)), vjust = .3, hjust = 1.1, color = "white", size = 3.5) +
  scale_x_continuous(labels = scales::dollar_format(scale = 1)) +
  labs(title = "Top 10 Schools that earned the least revenue in 2019",
       x = "Total Revenue",
       y = "Institution Name")
```

### What sports programs generate the most revenue for their schools?
```{r}
sports %>%
  group_by(sports) %>%
  summarise(total_revenue = sum(total_rev_menwomen, na.rm = TRUE)) %>%
  arrange(desc(total_revenue)) %>%
  head(10)

top_schools <- sports %>%
  group_by(institution_name) %>%
  summarise(school_total_revenue = sum(total_rev_menwomen, na.rm = TRUE)) %>%
  arrange(desc(school_total_revenue)) %>%
  head(10) %>%
  pull(institution_name)

# List the revenues of sports programs for each of these top 10 schools
sports_breakdown <- sports %>%
  filter(institution_name %in% top_schools, 
         !(sports %in% c("Rifle", "Weight lifting", "Archery", "Synchronized Swimming", "Diving", "Badminton", "Table Tennis", "Team Handball", "Rodeo", "Skiiing", "Track and Field, Indoor", "Track and Field, X-Country", "Track and Field, Outdoor"))) %>%
  group_by(institution_name, sports) %>%
  summarise(sport_revenue = sum(total_rev_menwomen, na.rm = TRUE)) %>%
  arrange(institution_name, desc(sport_revenue))

print(sports_breakdown)

# Let's assume your tibble is stored in a variable named `sports_revenue`
sports_breakdown %>%
  ggplot(aes(x = sports, y = sport_revenue, fill = institution_name)) +
  geom_bar(stat = "identity", position = "dodge") +  
  scale_y_continuous(labels = scales::dollar_format(scale = 1)) +  # Format the revenue in dollar format
  coord_flip() +
  labs(title = "Sports Revenue by Institution",
       x = "Sport",
       y = "Revenue") +
  theme_minimal() +
  theme(legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))
```

### Revenue by sports for UT Austin
```{r}
sports %>%
  select(year, institution_name, partic_men, partic_women, total_rev_menwomen, sports) %>%
  filter(!is.na(total_rev_menwomen) & institution_name == "The University of Texas at Austin" & year == 2019) %>%
  arrange(desc(total_rev_menwomen))
```

### What is the total revenue generated from college sports in Texas? 
```{r}
sports %>%
  select(state_cd, total_rev_menwomen) %>%
  group_by(state_cd = "TX") %>%
  summarise(total_rev_menwomen = sum(total_rev_menwomen, na.rm = TRUE))
```
The total revenue generated from college sports in Texas was sixty-nine billion two hundred ninety-one million six hundred eighty thousand forty-two (69291680042)

### Revenue by sports for Indiana University
```{r}
indianaUniversity <- sports %>%
  select(year, institution_name, sports, total_rev_menwomen) %>%
  filter(year == 2019, institution_name == "Indiana University-Bloomington") %>%
  arrange(desc(total_rev_menwomen)) %>%
  head(15) %>%
  ggplot(aes(total_rev_menwomen, sports, fill = sports)) +
  geom_bar(stat = "identity") +
  scale_x_continuous(labels = scales::dollar_format(scale = 1)) +
  labs(title = "Revenue by sports for Indiana University",
       x = "Sport",
       y = "Revenue") +
  theme_minimal() +
  theme(legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))

print(indianaUniversity)
```

### What Sport do women participate in the most in college?
```{r}
sports %>%
  select(sports, partic_women) %>%
  group_by(sports) %>%
  summarise(total_women = sum(partic_women, na.rm = TRUE)) %>%
  arrange(desc(total_women))
```

### What is the distribuion between men and womens in different sports in college?
```{r}
menWomenParicSports <- sports %>%
  select(sports, partic_men, partic_women) %>%
  group_by(sports) %>%
  summarise(
    total_women = sum(partic_women, na.rm = TRUE), 
    total_men = sum(partic_men, na.rm = TRUE)
  )

library(scales)

menWomenParicSports %>%
  ggplot(aes(x = sports)) +
  geom_bar(aes(y = total_men, fill = "Men"), stat = "identity") +
  geom_bar(aes(y = total_women, fill = "Women"), stat = "identity", position = "stack") +
  coord_flip() +
  labs(title = "Distribution of Men and Women in College Sports",
       x = "Sport", y = "Number of Participants") +
  theme_minimal() +
  scale_fill_manual(values = c("Men" = "blue", "Women" = "pink")) +
  scale_y_continuous(labels = comma)
```

### What is the revenue distrubtion for different sports by men and women? 
```{r}
totalRevMen <- sports %>%
  select(sports, rev_men) %>%
  group_by(sports) %>%
  summarise(total_rev_men = sum(rev_men, na.rm = TRUE)) %>%
  arrange(desc(total_rev_men))

totalRevWomen <- sports %>%
  select(sports, rev_women) %>%
  group_by(sports) %>%
  summarise(total_rev_women = sum(rev_women, na.rm = TRUE)) %>%
  arrange(desc(total_rev_women))

combined_data <- left_join(totalRevMen, totalRevWomen, by = "sports")

ggplot(data = combined_data, aes(x = reorder(sports, -total_rev_men))) +   # Ordering bars by total_rev_men
  geom_bar(aes(y = total_rev_men, fill = "Men"), stat = "identity", position = "dodge") +
  geom_bar(aes(y = total_rev_women, fill = "Women"), stat = "identity", position = "dodge") +
  labs(
    title = "Revenue Comparison: Men vs. Women in College Sports",
    y = "Total Revenue",
    x = "Sport",
    fill = "Gender"
  ) +
  theme_minimal() + 
  scale_y_continuous(labels = comma) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### What universities have the most male and female students?
```{r}
sports %>%
  select(institution_name, year, ef_female_count, ef_male_count, ef_total_count) %>%
  filter(ef_female_count > 17000 & ef_male_count > 17000 & year == 2019) %>%
  distinct() %>%
  arrange(desc(ef_total_count))
```

### What is the total male count vs women count across UT Austin?
```{r}
sports %>%
  filter(institution_name == "The University of Texas at Austin") %>%
  distinct(ef_male_count, year)
         
UTAustinEfMaleCount <- sports %>%
   filter(institution_name == "The University of Texas at Austin") %>%
  distinct(ef_male_count, year)

UTAustinEfFemaleCount <- sports %>%
   filter(institution_name == "The University of Texas at Austin") %>%
  distinct(ef_female_count, year)

combinedUTAustinMenWomen <- left_join(UTAustinEfMaleCount, UTAustinEfFemaleCount, by = "year")

library(ggplot2)
library(tidyr)

combinedUTAustinMenWomen %>%
  pivot_longer(cols = c(ef_male_count, ef_female_count), 
               names_to = "gender", 
               values_to = "count") %>%
  ggplot(aes(x=year, y=count, color=gender, group=gender)) +
  geom_line() +
  geom_point() +
  geom_text(aes(label=count), vjust=-0.5, size=3.5) +  # This line adds labels. You can adjust `vjust` for vertical position and `size` for text size
  labs(title = "UT Austin Attendance Over Time by Gender",
       x = "Year",
       y = "Number of Students",
       color = "Gender") +
  theme_minimal() +
  scale_color_manual(values = c("ef_male_count" = "blue", "ef_female_count" = "pink"))
```

### How has sports revenue changed over time for UT Austin?
```{r}
library(scales)
sports %>%
  filter(institution_name == "The University of Texas at Austin") %>%
  group_by(year) %>%
  summarize(total_revenue = sum(total_rev_menwomen, na.rm = TRUE)) %>%
  ggplot(aes(year, total_revenue)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = scales::dollar_format(scale = 1))
```

### What is football revenue overtime for the top 10 universities by total sports revenue?
```{r}
sports %>%
  group_by(sports, year) %>%
  summarise(total_revenue = sum(total_rev_menwomen, na.rm = TRUE)) %>%
  arrange(desc(total_revenue)) %>%
  head(10)

top_schools_year <- sports %>%
  group_by(institution_name) %>%
  summarise(school_total_revenue = sum(total_rev_menwomen, na.rm = TRUE)) %>%
  arrange(desc(school_total_revenue)) %>%
  head(10) %>%
  pull(institution_name)

# List the revenues of sports programs for each of these top 10 schools
sports_breakdown_year <- sports %>%
  filter(institution_name %in% top_schools_year, 
         !(sports %in% c("Rifle", "Weight lifting", "Archery", "Synchronized Swimming", "Diving", "Badminton", "Table Tennis", "Team Handball", "Rodeo", "Skiiing", "Track and Field, Indoor", "Track and Field, X-Country", "Track and Field, Outdoor"))) %>%
  group_by(institution_name, sports, year) %>%
  summarise(sport_revenue = sum(total_rev_menwomen, na.rm = TRUE)) %>%
  arrange(institution_name, desc(sport_revenue))

print(sports_breakdown)

sports_breakdown_year %>%
  filter(sports == "Football") %>%
  group_by(year)

sports_breakdown_year %>%
  filter(sports == "Football") %>%
  ggplot(aes(x=year, y=sport_revenue, color=institution_name)) +
  geom_line(aes(group=institution_name)) +
  geom_point() +
  facet_wrap(~institution_name, scales = "free_y") + 
  labs(title="Football Revenue by Year for Different Schools",
       x="Year",
       y="Revenue ($)",
       color="School") +
  scale_y_continuous(labels = scales::dollar_format(label = 1)) +
  theme_minimal()
```

### What states have the schools that generate the highest revenue?
```{r}
top_states_rev <- sports %>%
  group_by(state_cd) %>%
  summarise(total_revenue = sum(total_rev_menwomen, na.rm = TRUE)) %>%
  arrange(desc(total_revenue))

print(top_states_rev)

# Install & Load the required packages
options(repos = c(CRAN = "https://cloud.r-project.org/"))
install.packages(c("sf", "viridis"))
library(sf)
library(ggplot2)
library(viridis)
library(maps)
library(scales)

# Add a column of state names to top_states_rev based on abbreviations
top_states_rev$state_name <- state.name[match(top_states_rev$state_cd, state.abb)]

convert_state_code <- function(code) {
  state_lookup <- data.frame(code = state.abb, name = tolower(state.name))
  return(state_lookup$name[state_lookup$code == code])
}

top_states_rev$state_name <- sapply(top_states_rev$state_cd, convert_state_code)

# Get US Map Data
us_states <- maps::map("state", plot = FALSE, fill = TRUE) %>% 
  st_as_sf()

top_states_rev$state_name <- as.character(top_states_rev$state_name)

# Merge the map data with your `top_states` data
merged_data <- left_join(us_states, top_states_rev, by = c("ID" = "state_name"))

# Custom function to format numbers into "M" for millions
format_millions <- function(x) {
  return(scales::dollar_format(scale = 1e-6, suffix = "M")(x))
}

breaks_to_use <- scales::pretty_breaks(n = 5)(range(merged_data$total_revenue))

ggplot(data = merged_data) + 
  geom_sf(aes(fill = total_revenue)) + 
  scale_fill_viridis(name = "Total Revenue", option = "C", direction = -1,
                     labels = scales::label_number(scale = 1e-6, suffix = 'M')) +
  theme_light() +
  labs(title = "Total Revenue by State for College Sports") + 
  theme(legend.position = "bottom", 
        legend.key.size = unit(1.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.ticks = element_blank(), 
        axis.text = element_blank(),
        panel.border = element_blank())
```

# Inferential Statistics

### Difference in Participation Rates between Genders
```{r}
t.test(sports$partic_men, sports$partic_women)
```
Based on the t-test results, we can conclude that there is a statistically significant difference between the means of participation rates between men and women. Specifically, the mean for partic_men is higher than that for partic_women, with a difference estimated to be between 9.88 and 10.42. This difference is highly significant, given the extremely small p-value.

### Correlation between Revenue and Expenditure for Men's Sports
```{r}
cor.test(sports$total_exp_menwomen, sports$exp_men)
```
There's a very strong positive linear relationship between total_exp_menwomen and exp_men. This means as expenditures for men increase, the total expenditures for both men and women also tend to increase, and vice versa.

The correlation coefficient of approximately 0.9846 is statistically significant with a p-value effectively at 0. This suggests that it's highly unlikely that such a strong correlation occurred by chance alone.

### Difference in revenue between men and womens sports
```{r}
t.test(sports$rev_men, sports$rev_women, paired = TRUE)
```
Based on the t-test results, we can say that there's a statistically significant difference between the revenues of men (rev_men) and women (rev_women) in sports. Specifically, the revenue for men appears to be higher than that for women by an average of approximately $135206.3.

### Association between School Classification and Sports Offered
```{r}
table_data <- table(sports$classification_name, sports$sports)
chisq.test(table_data)
```
The result indicates that different classifications of schools likely have different distributions of sports offered.