---
title: "Alone"
author: "Braden Buchanan"
date: "2023-09-06"
output: 
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)ls
```

# Load Alone Tidy Tuesday Data
```{r}
survivalists <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-24/survivalists.csv')
loadouts <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-24/loadouts.csv')
episodes <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-24/episodes.csv')
seasons <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-24/seasons.csv')
```

Load packages
```{r}
library(tidyverse)
```

# Descriptive Statistics

### Where are the different participants from in the country?
```{r}
survivalists %>%
  group_by(country) %>%
  count()
```
Majority of participants are from the United States. In total across 9 seasons, 79/94 from the United States, followed by 13/94 from Canada.

### What states are most US participants from?
```{r}
survivalists %>%
  group_by(state) %>%
  count() %>%
  arrange(desc(n))
```

### What country are most winners from?
```{r}
survivalists %>%
  select(season, country, result, days_lasted, profession) %>%
  filter(result == 1) %>%
  group_by(country)
```

### Male vs female winners
```{r}
survivalists %>%
  filter(result == "1") %>%
  group_by(season, gender, days_lasted) %>%
  summarise(winner_count = n())
```
Over 9 seasons of Alone, only males have came in first place. 

### How many women came in 2nd place and avg days lasted?
```{r}
survivalists %>%
  filter(result == 2 & gender == "Female") %>%
  group_by(season, gender, days_lasted) %>%
  summarize(second_place = n())

survivalists %>%
  select(gender, days_lasted, result) %>%
  filter(result == 2 & gender == "Female") %>%
  mutate(avg_days_lasted = mean(days_lasted))
```
4x Women came in second place. On average, the 2nd place women last 80.8 days

### What were the main reasons for participants to tap out?
```{r}
survivalists %>%
  select(season, reason_tapped_out, result, gender) %>%
  filter(!is.na(reason_tapped_out)) %>%
  mutate(reason_tapped_out = ifelse(str_detect(reason_tapped_out, "Starvation"), "Starvation", reason_tapped_out)) %>%
  count(reason_tapped_out) %>%
  arrange(desc(n))
```
Most participants tap out due to starvation or missing their family.

### Main reason for 2nd place for tapping out?
```{r}
library(stringr)

survivalists %>%
  select(season, reason_tapped_out, result, gender) %>%
  filter(!is.na(reason_tapped_out), result == 2) %>%
  mutate(reason_tapped_out = ifelse(str_detect(reason_tapped_out, "Starvation"), "Starvation", reason_tapped_out)) %>%
  count(reason_tapped_out) %>%
  arrange(desc(n))
```
Main reasons participants in 2nd place tapped out was due to starvation.

### Most common equipment participants had
```{r}
mostCommonLoadouts <- loadouts %>%
  select(item_number, item_detailed, item) %>%
  count(item) %>%
  mutate(item = fct_reorder(item, n)) %>%
  arrange(desc(n))

ggplot(mostCommonLoadouts, aes(x = item, y = n, fill = item)) +
  geom_bar(stat="identity") +
    geom_text(aes(label=n), position=position_dodge(width=0.9), hjust=-0.3) +
  coord_flip() + 
  labs(x = "Item", 
       y = "Number of Items") + 
  theme(legend.position = "none")
```
Pots, fishing gear, and sleeping bags were the most common loadouts for participants.

# Inferential Statistics
Use a t-test to see if gender significantly affects the number of days lasted.
```{r}
# Filtering data to exclude rows where 'days_lasted' is NA
filtered_data <- survivalists %>%
  filter(!is.na(days_lasted))

t_test_result <- t.test(days_lasted ~ gender, data = filtered_data)

print(t_test_result)
```
Men and women seem to last different amounts of time on the show, with women lasting longer on average. However, we're not 100% sure if this difference is because of gender or just happened by chance. We'd need more evidence or a stronger result to be certain.

### Is age correlated with success in the show?
```{r}
# Compute the correlation
correlation <- cor(survivalists$age, survivalists$days_lasted, use = "complete.obs")

print(correlation)


# Compute the correlation for age and result
correlation_result <- cor(survivalists$age, survivalists$result, use = "complete.obs")

print(correlation_result)

# Correlation test for age and days_lasted
test_result <- cor.test(survivalists$age, survivalists$days_lasted, use = "complete.obs")

print(test_result)
```
Based on the data and tests conducted, age doesn't seem to have a significant impact on how long participants last in the show or on their result rankings.

There is a weak positive correlation of 0.0662 between age and the number of days participants lasted in the show. However, this correlation is not statistically significant with a p-value of 0.5259.

There's also a weak negative correlation of -0.0419 between age and the result, but the significance of this relationship wasn't tested in the provided outputs.

### Are survivalists from certain professions more likely to succeed?
```{r}
contingency_table <- table(survivalists$profession, survivalists$result)
chi_squared_test <- chisq.test(contingency_table)
print(chi_squared_test)
```
The data doesn't provide strong evidence to suggest that the profession of survivalists is associated with their success on the show.

The p-value is 0.4184, which is greater than 0.05. This suggests that, based on the data we don't have enough evidence to reject the null hypothesis. 