---
title: "Alone"
author: "Braden Buchanan"
date: "2023-09-06"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load Alone Tidy Tuesday Data
```{r}
survivalists <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-24/survivalists.csv')
loadouts <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-24/loadouts.csv')
episodes <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-24/episodes.csv')
seasons <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-24/seasons.csv')
```

Load packages
```{r}
library(tidyverse)
```

#Descriptive Statistics

###Where are the different participants from in the country?
```{r}
survivalists %>%
  group_by(country) %>%
  count() %>%
```
Majority of participants are from the United States. In total across 9 seasons, 79/94 from the United States, followed by 13/94 from Canada.

###What states are most US participants from?
```{r}
survivalists %>%
  group_by(state) %>%
  count() %>%
  arrange(desc(n))
```


### Male vs female winners
```{r}
survivalists %>%
  filter(result == "1") %>%
  group_by(season, gender, days_lasted) %>%
  summarise(winner_count = n())
```
Over 9 seasons of Alone, only males have came in first place. 

###How many women came in 2nd place?
```{r}
survivalists %>%
  filter(result == 2 & gender == "Female") %>%
  group_by(season, gender, days_lasted) %>%
  summarize(second_place = n())
```


