---
title: "London Marathon"
author: "Braden Buchanan"
date: "2024-01-05"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
winners <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-04-25/winners.csv')
london_marathon <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-04-25/london_marathon.csv')

theme_set(theme_minimal())
```

Exploring Data
```{r}
view(winners)
view(london_marathon)
```

Change in applications over time 
```{r}
london_marathon %>%
  select(Year, Applicants) %>%
  ggplot(aes(Year, Applicants)) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "London Marathon applications over time")
```

What is the acceptance rate for the London marathon?
```{r}
london_marathon %>%
  mutate(acceptance_rate = Accepted/Applicants)%>%
  select(Year, Applicants, acceptance_rate) %>%
  ggplot(aes(Year, acceptance_rate)) +
  geom_line() +
  labs(title = "London Marathon Acceptance Rate Over Time",
       y = "Acceptance Rate")
```

What year had the least amount of finishers in the London Marathon?
```{r}
london_marathon %>%
  mutate(percent_finished = Finishers/Starters) %>%
  select(Year, Starters, Finishers, percent_finished) %>%
  arrange(percent_finished)
```

What is the fastest marathon time for categories men and women?
```{r}
winners %>%
  filter(Category %in% c("Men", "Women")) %>%
  group_by(Category) %>%
  arrange(Time)
```

How have the London Marathon Winner's time changed over time?
```{r}
winners %>%
  filter(Category %in% c("Men", "Women")) %>%
  select(Year, Time, Category) %>%
  ggplot(aes(Year, Time, color = Category)) +
  geom_point() +
  labs(title = "How London Marathon Winners Times have changed over time",
       subtitle = "Categories Men & Women")
```

What Nationality wins the most in the London Marathon?
```{r}
winners %>%
  filter(Category %in% c("Men", "Women")) %>%
  group_by(Nationality) %>%
  count(Nationality) %>%
  arrange(desc(n))


winners %>%
  filter(Category %in% c("Men", "Women")) %>%
  count(Nationality) %>%
  mutate(Nationality = fct_reorder(Nationality, n)) %>%
  ggplot(aes(x = Nationality, y = n, fill = Nationality)) +
  geom_col() +
  coord_flip() +
  labs(title = "Nationality that wins the most in the London Marathon",
       subtitle = "Categories Men & Women",
       y = "Count") +
  theme(legend.position = "none")
```

What athlete has won the London Marathon the most?
```{r}
winners %>%
  group_by(Athlete) %>%
  count(Athlete) %>%
  arrange(desc(n))
```

