---
title: "Company reputation poll"
author: "Braden Buchanan"
date: "2023-12-17"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages
```{r}
library(tidytuesdayR)
library(tidyverse)
```

# Load Data
```{r}
tuesdata <- tidytuesdayR::tt_load("2022-05-31")

reputation <- tuesdata$reputation
poll <- tuesdata$poll
```

# Examine Reputation Data
```{r}
summary(reputation)
head(reputation)
```

### How many companies are in this list?
Reputation list accounts for 100 companies
```{r}
reputation %>%
  select(company) %>%
  unique() %>%
  count()
```

### What industries are the most represented in this data?
```{r}
mostCommonIndustries <- reputation %>%
  select(company, industry) %>%
  unique() %>%
  count(industry) %>%
  arrange(desc(n))

mostCommonIndustries

ggplot(mostCommonIndustries, aes(fct_reorder(industry, n), n, fill=industry)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = n), hjust = 1.5, vjust = .4, size = 2.5) +
  coord_flip() + 
  theme(legend.position = "none") +
  ggtitle("Most common industries in Reputation data set") +
  ylab("Count") + 
  xlab("Industry")
```

### What Tech companies had the highest reputation on average?
Samsung, Microsoft, and IBM are the top 3 tech companies by average score. Tiktok, Facebook, and Twitter are the lowest 3 tech companies
```{r}
reputation %>%
  select(company, industry, name, score, rank) %>%
  filter(industry == "Tech") %>%
  group_by(company) %>%
  summarize(avg_score = mean(score), .groups = "drop") %>%
  arrange(desc(avg_score))
```

### What tech company has the highest trust rating?
```{r}
techTrustRatings <- reputation %>%
  select(company, industry, name, score, rank) %>%
  filter(industry == "Tech", name == "TRUST") %>%
  group_by(company) %>%
  arrange(desc(score))

techTrustRatings

ggplot(techTrustRatings, aes(fct_reorder(company, score), score, fill = company)) +
  geom_bar(stat = "identity") +
  coord_flip() + 
  theme(legend.position = "none")
```

